Metabolic Effects of PCSK9 Inhibition With Evolocumab in Subjects with Elevated Lp(a)

### 2019-12-13 manuscript.Rmd
story Lp(a) and TG metabolism

---
title: "Metabolic Effects of PCSK9 Inhibition With Evolocumab in subjects with elevated Lp(a)"
output: word_document
bibliography: /home/xiang/Documents/Bibliography/lpa.bib
csl: circulation.csl
---

@@

Background: 
Epidemiological studies show that subjects with elevated lipoprotein(a) [Lp(a)] have a higher risk for developing cardiovascular diseases. 
Inhibition of proprotein convertase subtilisin/kexin type 9 (PCSK9) is one of the limited options to reduce Lp(a). 
In this study, we aimed to investigate metabolic effects of PCSK9 inhibition with evolocumab in subjects with elevated Lp(a).

Methods:
We performed metabolic profiling of plasma samples derived from 30 individuals (placebo N = 14; evolocumab N = 16). 
By applying Bayesian lognormal regression, we assessed the effects of evolocumab treatment on circulating metabolites such as lipoproteins and amino acids. 
In order to separate Lp(a) content from the LDL measurements, we applied a Gaussian mixture model to estimate the proportions of LDL particles with and without apolipoprotein(a) [apo(a)].

Results:
We identified that evolocumab treatment for 16 weeks resulted in 17% (95% credible interval: 8% to 26%) reduction of circulating Lp(a), coupled with substantial reduction of VLDL, IDL and LDL particles as well as their lipid contents.  
We also identified that about 24% [0.2% 47%] of large LDL particles were with apo(a). 
We did not observe changes in the proportion of LDL particles with apo(a) after treatment of evolocumab.

Conclusions:
Our analysis suggested that alternative pathways beyond LDL receptor mediated clearance may be involved in Lp(a) reduction by PCSK9 inhibition.

@@

lipoprotein(a) [Lp(a)], synthesized in liver, is a low-density lipoprotein (LDL)-like particle with an apolipoprotein(a) [apo(a)] tail. 
Similar to other lipoproteins, Lp(a) is cleared from the circulation primarily by the liver [@Cain2005]. 
Epidemiological studies showed that subjects with elevated Lp(a) had increasing risk of developing cardiovascular disease [@ODonoghue2019]. 
Unlike LDL-cholesterol lowering, limited therapeutic options are available to reduce plasma Lp(a) concentration [@VanCapelleveen2016]. 
One of the agents that are capable to lower circulating Lp(a) levels is proprotein convertase subtilisin/kexin type 9 (PCSK9) monoclonal antibody, which prevents extracellular PCSK9 from binding to the LDL receptor, and results in supraphysiological abundance of LDL receptors [@Raal2014; @Raal2016].
The excess number of LDL receptors on the surface of hepatocytes were hypothesized to be responsible for the enhanced Lp(a) catabolism [@Romagnuolo2015; @Raal2016]. 
However, the above hypothesis does not concur with the effect of statins therapy, which significantly increases hepatic LDL receptor abundances but also increases plasma Lp(a) levels [@Tsimikas2019]. 
The same hypothesis was also challenged by the observation that LDL receptor knock out mice had similar Lp(a) catabolism as the wild type mice [@Cain2005]. 
Therefore, what roles LDL-receptor plays in Lp(a) catabolism has not been elucidated and remains a matter of debate [@Reyes-Soffer2017; @Watts2018; @Stoekenbroek2018]. 
In addition to the LDL receptor, very low density lipoprotein (VLDL) receptor and scavenger receptor class B type I (SR-BI) were also found to act as potential receptor for Lp(a), indicating more than one pathway is likely to be involved in Lp(a) reduction by PCKS9 inhibition [@Argraves1997; @Yang2013].
As a cost-effective approach, nuclear magnetic resonance (NMR) metabolomics simultaneously quantifies hundreds of circulating biomarkers and provides insights on multiple pathways [@Soininen2009].
In fact, NMR metabolomics has been used to study metabolic effects that were associated with  *LPA* loci, as well as genetic inhibition of PCSK9 [@Kettunen2016; @Sliz2018]. 
However, genetic inhibition of PCSK9 is different from inhibition with monoclonal antibody, and there is lack of studies on metabolic effects of PCSK9 inhibition with monoclonal antibody. 
In the current study, we aim to study metabolic changes corresponding to Lp(a) reduction in subjects with elevated Lp(a) after treatment of PCSK9 monoclonal antibody (evolocumab).
To this end, we performed NMR metabolomics of plasma samples derived from 30 healthy participants with elevated Lp(a) levels (placebo 14; evolocumab 16). 
Since Lp(a) and LDL significantly overlap in density, protein and lipid composition, LDL measurements made by current technologies contain the content of Lp(a) [@Yeang2015]. 
To address this issue, in this study we developed a statistical model to distinguish these two lipoproteins.

@@

Patients were found to be eligible for enrollment in the ANITSCHKOW trial when they were $\geq$ 50 years of age, had a fasting LDL-C of $\geq$ 2.6 mmol/L (100mg/dL), and an Lp(a) of $\geq$ 125nmol/L (50mg/dL). 
At time of randomization, study subjects who were on lipid lowering treatment had to be on a stable dosing regimen for at least 8 weeks. 
Exclusion criteria were among others, diabetes mellitus and a cardiovascular event within three months prior to randomization. 
The complete list of eligibility criteria was reported previously [@Stiekema2018].

@@

Blood samples were obtained at time of randomization and sixteen weeks after treatment. 
Patients were fasting for $\geq$ 9 hours for both blood withdrawals.

Lipoprotein(a) levels were measured using an isoform-independent immunoturbidometric assay (Polymedco, Cortlandt Manor, NY, USA) and reported in nmol/L.

@@

To assess the effect of evolocumab on levels of a circulating metabolite, we ran lognormal regression analysis. 
The outcome variable ($y$) was concentration of a metabolite. 
Two predictor variables were in the regression model: 1) evolocumab treatment (evolocumab, $T_i$ = 1; placebo $T_i = 0$); 
2) metabolite concentration at time of randomization ($x$).
The baseline concentrations were centered and scaled so that the mean is 0 and standard deviation is 1.
The lognormal distribution was chosen to model the outcome variable because its values were positive continuous. 
Due to missing observations in outcome variables, we applied the Bayesian approach to handle the missing data. 
There are two types of missing values, which are as follows: 
(1) when the concentration of a metabolite is below the limit of detection, 
or (2) when values were rejected by the automatic sample and measurement quality control procedure in the Nightingale pipeline. 
All the missing observations were assumed missing at random and treated as parameters.
Values were randomly drawn from a lognormal distribution. 
Regarding the missing values that were below the limit of detection, the imputed values were constrained between zero and the minimal observed value. 
We fitted the model by running Hamiltonian Markov Chain Monte Carlo in the program Stan (version 2.18.3).
We ran four Markov chains with 2000 iterations in each chain. 
Results were presented with the posterior mean of lognormal regression coefficient with 95% credible interval (CI). 
The lognormal regression coefficient can be interpreted as the expected proportional change in the metabolite concentration.

### Separating Lp(a) content from LDL measures

Since Lp(a) and LDL significantly overlap in density, as well as protein and lipid composition, LDL measures made by current technologies contain the content of Lp(a). 
In this study, we applied the Gaussian mixture model to address this issue. 
Our statistical model was inspired by the biological relationship between Lp(a) and LDL. 
Lp(a) is often considered to comprise a LDL particle and covalently bound apo(a). 
The diameter of Lp(a) is about 25 nm [@BERG1963]. 
Based on the above knowledge, we assumed that the pool of large LDL particle (average diameter 25.5 nm) is a mixture of LDL particles with and without apo(a). 
The outcome variable ($y$) was the particle concentration of large LDL, and described by a mixture of two normal distributions with a proportion parameter $\lambda$.

$$P(y|\lambda, \mu, \sigma) = \prod_{i = 1}^{N}(\lambda \times N(\mu_1,\sigma_1) + (1 - \lambda) \times N(\mu_2, \sigma_2))$$

The parameter $\mu_1$ and $\sigma_1$ were estimated based on the particle concentration of Lp(a), which was measured by using the isoform-independent immunoturbidometric assay. 
The parameter $\mu_2$ was modeled as $\mu_2[i] = \alpha_L + \beta_{S,L} \times S[i] + \beta_{E,L} \times E[i] + (\beta_L + \beta_{P,L} \times P[i]) \times V[i]$.

There were four covariates $V$, $S$, $E$ and $P$. 
$V = 0$ referred the measurement was made at time of randomization, whereas $V = 1$ refereed to sixteen weeks after treatment. 
Because lipid-lowering medication such as statins, ezetimibe as well as evolocumab has causal effects on these LDL particle concentrations, we included three covariates (statin usage: $S = 1$; Ezetimibe usage $E = 1$; Evolocumab usage $P = 1$). 
Regarding parameter $\lambda$,  proportion of LDL particles with apo(a) tail, we set the beta distribution $Beta(1,1)$ as its prior distributions.

@@

The current study included 30 patients (evolocumab, n = 16; placebo, n = 14). 
Baseline characteristics of the two groups were comparable and provided in Table 1.

@@

Evolocumab treatment for 16 weeks resulted in mean 17% (95% credible interval: 8% to 26%) reduction in Lp(a) compared with placebo.

To identify the metabolic effects corresponding to the Lp(a) reduction by PCSK9 inhibition, we performed NMR metabolomics covering metabolic pathways such as lipoproteins, fatty acids, amino acids and glycolysis. 
We observed that evolocumab treatment resulted in substantial reduction in particle concentration of extremely large (81% [51% 100%]), very large (89% [69% 100%]), large (61% [36%, 86%]), medium (50% [36%, 63%]), small (39% [31% 46%]) and very small VLDL (47% [41% 54%]) compared with the placebo group.
We also observed that evolocumab treatment resulted in particle concentration reduction in IDL (53% [45%, 61%]), large (57% [49% 66%]), medium (59% [50%, 6%]) and small LDL (55% [47%, 64%]) compared with the placebo group. 
In addition, we observed that evolocumab treatment resulted in decreased concentrations of very large HDL particles (25% [4% 46%]), and increased concentrations of medium HDL particles (13% [4% 23%]) (Figure 1).

@@

Similar to the lipoprotein particle concentration profiles, we observed that evolocumab treatment resulted in substantial reduction of esterified cholesterol (CE), triglyceride (TG), free cholesterol (FC) and phospholipid (PL) in VLDL, IDL and LDL as well as very large HDL, compared with the placebo group (Figure 2).

@@

We observed no difference in concentrations of other metabolites covering amino acids, fluid balance, glycolysis and ketone bodies.

## Separating Lp(a) content from the LDL measures

Lp(a) is thought to be comprised of a LDL particle and a covalently bound apo(a). 
Because Lp(a) and LDL significantly overlap in density, as well as protein and lipid composition, LDL measures contain the content of Lp(a). 
In this study, we focused on the large LDL particle, because its diameter (25.5 nm) is close to the diameter of Lp(a). 
At baseline, we identified that on average 24% ([0.2%, 47%]) large LDL particles were with apo(a) (Figure 4).

Sixteen weeks after treatment of evolocumab, on average 12% (0.02%, 36%) large LDL particles were with apo(a) in the placebo group. 
In the treatment group, the average proportion of large LDL particles with apo(a) was 9% (0.0008%, 31%).
We did not observe that the proportions of large LDL particles with apo(a) were different between the treatment and placebo group (-3% [-36% 23%]).

@@

The major finding of this work is that evolocumab treatment in patients with high Lp(a) massively decreases large VLDL in addition to the well known effect on LDL. 
In the current work, evolocumab treatment resulted in on average 17% reduction in plasma Lp(a) concentration compared with placebo. 
Our detailed metabolomic profiling revealed that Lp(a) reduction by PCSK9 inhibition corresponded to substantial reduction in concentrations of VLDL, IDL and LDL particles as well as their lipid and fatty acid contents. 
Our observation is consistent with the previous study, in which evolocumab treatment not only reduced concentrations of LDL particles, but also substantially reduced triglyceride-rich lipoprotein particles [@Toth2018]. 
Since a PCSK9 monoclonal antibody such as evolocumab reduces the degradation of the LDL receptor as well as the VLDL receptor, the very low concentration of VLDL particles after the treatment may be ascribed to the increased number of VLDL receptors [@Poirier2008]. 
The upregulation of VLDL receptors can enhance uptake of VLDLs but also Lp(a) from the circulation [@Argraves1997]. 
In addition, our analysis also showed that treatment of evolocumab lowered concentrations of very large HDL particles. 
Very large HDL is cholesterol enriched and mainly captured by the SR-B1 present in the liver [@Acton1996]. 
Similar to the VLDL receptor, SR-B1 has also been shown to bind Lp(a), and its upregulation will probably enhance the Lp(a) catabolism [@Yang2013].

In our analysis, we observed a large discordance in Lp(a) and LDL reduction in response to evolocumab. 
This discordance was suggested by a recent study to be related to alternative pathways of Lp(a) reduction beyond LDL receptor mediated clearance [@Shapiro2019]. 
Lp(a) and LDL significantly overlap in density, as well as protein and lipid composition. 
As a result, current technologies including the NMR metabolomics used in this study cannot distinguish these two lipoproteins, and include the content of Lp(a) in the LDL measurements. 
We think that it is important to separate the Lp(a) content from the LDL measurements in order to better understand the discordance between these two lipoproteins in response to evolocumab. 
To this end, we applied a Gaussian mixture model based on the assumption that Lp(a) was comprised of a LDL particle and a covalently bound apo(a). 
We identified that on average about 24% of large LDL particle had apo(a) tail. 
We did not observe changes in proportions of large LDL particle with apo(a) tail due to evolocumab treatment.

@@

Strengths and limitations of our study warrant consideration. 
Our study provided the first NMR metabolomics data for a PCSK9 inhibition trial. 
The detailed metabolomic profiling revealed metabolic effects of the PCSK9 inhibitor, and improved our understanding of this lipid-lowering therapy. 
Moreover, in this study we developed a statistical method to separate Lp(a) content from the LDL measurements. 
However, due to the heterogeneity of Lp(a) and limited number of participants, our estimation of the proportions of LDL with apo(a) had very large uncertainty. 
Applying our statistical model with a larger number of participants as well as validation in the laboratory are warranted in the future studies.

In conclusion, our NMR metabolomic profiling revealed that Lp(a) reduction by PCSK9 inhibition corresponded to substantial reduction in VLDL, IDL and LDL particles as well as their lipid contents. 
Our analysis suggested that alternative pathways beyond LDL receptor mediated clearance may be involved in Lp(a) reduction by PCSK9 inhibition.

>>>>

---
title: "Metabolic Effects of PCSK9 Inhibition With Evolocumab in Subjects with Elevated Lp(a)"
output: word_document
bibliography: /home/xiang/Documents/Bibliography/lpa.bib
csl: journal-of-lipid-research.csl
---

@@

Background: 
Epidemiological studies show that subjects with elevated lipoprotein(a) [Lp(a)] have a higher risk for developing cardiovascular diseases. 
Inhibition of proprotein convertase subtilisin/kexin type 9 (PCSK9) is one of the limited options to lower Lp(a). 
However, effects of PCSK9 inhibition with evolocumab on circulating metabolites such as lipoprotein subclasses, amino acids and fatty acids remain to be characterized.

Methods:
In this study, we performed nuclear magnetic resonance (NMR) metabolomics on plasma samples derived from 30 individuals (placebo N = 14; evolocumab N = 16). 
We assessed the effects of evolocumab treatment on circulating metabolites by running lognormal regression analyses. 
Subsequently, we assessed the interrelationship between Lp(a) and 14 lipoprotein subclasses in response to treatment of evolocumab, by running multilevel multivariate regression analyses.

Results:
On average, evolocumab treatment for 16 weeks resulted in a 17% (95% credible interval: 8% to 26%) reduction of circulating Lp(a), coupled with substantial reduction of VLDL, IDL and LDL particles as well as their lipid contents. 
Although, we did not observe correlations between Lp(a) lowering and reduction in other 14 lipoprotein subclasses, we identified that increasing concentrations of baseline Lp(a) were associated with larger reduction in triglyceride-rich VLDL particles after evolocumab treatment.

Conclusions:
Inhibition of PCSK9 with evolocumab massively decreased VLDL in addition to the well known effect on LDL. 
The substantial reduction in VLDLs after evolocumab treatment indicates enhanced clearance of VLDL particles in patients with high Lp(a).

@@

lipoprotein(a) [Lp(a)], synthesized in the liver, is a low-density lipoprotein (LDL)-like particle with an apolipoprotein(a) [apo(a)] tail. 
Similar to other lipoproteins, Lp(a) is cleared from the circulation primarily by the liver [@Cain2005]. 
Epidemiological studies showed that subjects with elevated Lp(a) had increasing risk of developing cardiovascular disease [@ODonoghue2019]. 
The biological mechanism underlying the cardiovascular risk of Lp(a) remains to be elucidated. 
One mechanism is that Lp(a) may induce macrophage apoptosis and plaque necrosis through activation of a pathway involving CD36 (Cluster of differentiation 36) and TLR2 (Toll-like receptor 2) [@Seimon2010]. 
Unlike LDL-cholesterol lowering, limited therapeutic options are available to reduce Lp(a) in plasma [@VanCapelleveen2016]. 
One of the agents that are capable to lower circulating Lp(a) levels is proprotein convertase subtilisin/kexin type 9 (PCSK9) monoclonal antibody, which prevents extracellular PCSK9 from binding to the LDL receptor, and results in supraphysiological abundance of LDL receptors [@Raal2014; @Raal2016].
The excess number of LDL receptors on the surface of hepatocytes were hypothesized to be responsible for the enhanced Lp(a) catabolism [@Romagnuolo2015; @Raal2016]. 
However, the above hypothesis does not concur with the effect of statin therapy, which significantly increases hepatic LDL receptor abundances but also increases plasma Lp(a) levels [@Tsimikas2019]. 
The same hypothesis was also challenged by the observation that LDL receptor knock out mice had similar Lp(a) catabolism compared to wild type mice [@Cain2005]. 
Hence, the exact role of the LDL-receptor in Lp(a) catabolism has not been elucidated and remains a matter of debate [@Reyes-Soffer2017; @Watts2018; @Stoekenbroek2018]. 
The LDL receptor is not the only target of PCSK9. 
PCSK9 also induces the degradation of the VLDL (very low density lipoprotein) receptor, ApoER2 (apolipoprotein E receptor 2), and CD36 [@Poirier2008; @Demers2015]. 
Interestingly, the VLDL receptor and CD36 are receptors for triglyceride-rich lipoproteins (TRLs) as well as Lp(a), indicating a possible interrelationship between Lp(a) and TRL metabolism [@Argraves1997; @Seimon2010; @Dijk2018]. 
The metabolic link between Lp(a) and TRL metabolism was confirmed by a recent NMR (nuclear magnetic resonance) metabolomics study, in which a Lp(a)-raising genotype was identified to associate with decreasing concentrations of triglyceride-rich VLDL particles [@Kettunen2016]. 
However, the mechanism underlying this inverse relationship is unknown. 
As a cost-effective approach, NMR metabolomics simultaneously quantifies hundreds of circulating biomarkers and provides insights on multiple metabolic pathways [@Soininen2009].
A NMR metabolomics study showed that genetic inhibition of PCSK9 and statin therapy had similar effects on circulating metabolites [@Sliz2018]. 
However, genetic inhibition of PCSK9 is different from inhibition with monoclonal antibody since genetic inhibition includes intracellular functions of PCSK9. 
Furthermore, although PCSK9 monoclonal antibody is known for lowering LDL cholesterol and Lp(a), the relationship between Lp(a) lowering and metabolism of other lipoproteins has not been studied.  
In the current study, we analyzed NMR metabolomics of plasma derived from 30 subjects with elevated Lp(a) (placebo 14; evolocumab 16). 
We first ran lognormal regression analyses to identify metabolic changes corresponding to evolocumab treatment.  
Next, we developed a multilevel multivariate model to assess the relationship between Lp(a) lowering and metabolism of 14 lipoprotein subclasses.

@@

Patients were found to be eligible for enrollment in the ANITSCHKOW trial when they were $\geq$ 50 years of age, had a fasting LDL-C of $\geq$ 2.6 mmol/L (100mg/dL), and an Lp(a) of $\geq$ 125nmol/L (50mg/dL). 
At the time of randomization, study subjects who were on lipid lowering treatment had to be on a stable dosing regimen for at least 8 weeks. 
Exclusion criteria were among others, diabetes mellitus and a cardiovascular event within three months prior to randomization. 
The complete list of eligibility criteria was reported previously [@Stiekema2018].

## Biochemical measurements

Blood samples were obtained at time of randomization and 16 weeks after treatment. Patients were fasting for $\geq$ 9 hours for both blood withdrawals. 
Total cholesterol, high-density lipoprotein (HDL) cholesterol, triglycerides, and apolipoproteinB-100 (ApoB-100) were measured by commercially available kits at the Medpace core lab (Medpace Reference Laboratories, Leuven, Belgium). 
Low-density lipoprotein (LDL) cholesterol was calculated using the Friedewald formula. 
Ultracentrifugation-determined LDL-C was measured and reported if calculated LDL cholesterol was below 40mg/dL, or triglycerides were above 400mg/dL. 
Lipoprotein(a) levels were measured using an isoform-independent immunoturbidometric assay (Polymedco, Cortlandt Manor, NY, USA) and reported in nmol/L.

@@

@@

To assess the effect of evolocumab on levels of a circulating metabolite, we ran lognormal regression analysis. 
The outcome variable ($y$) was the concentration of a metabolite. 
Two predictor variables were in the regression model: 1) evolocumab treatment (evolocumab, $T_i$ = 1; placebo $T_i = 0$); 
2) metabolite concentration at time of randomization ($x$).
The baseline concentrations were centered and scaled so that the mean is 0 and standard deviation is 1.
The lognormal distribution was chosen to model the outcome variable because its values were positive continuous. 
Due to missing observations in outcome variables, we applied the Bayesian approach to handle missing data. 
There were two types of missing values: 
(1) when the concentration of a metabolite is below the limit of detection, 
or (2) when values were rejected by the automatic sample and measurement quality control procedure in the Nightingale pipeline. 
All the missing observations were assumed missing at random and treated as parameters.
Values were randomly drawn from a lognormal distribution. 
Regarding the missing values that were below the limit of detection, the imputed values were constrained between zero and the minimal observed value. 
We fitted the model by running Hamiltonian Markov Chain Monte Carlo in the program Stan (version 2.18.3).
We ran four Markov chains with 2000 iterations in each chain. 
Results were presented with the posterior mean with 95% credible interval (CI).

### Multilevel multivariate model

To assess the relationship between Lp(a) lowering and metabolism of 14 lipoprotein subclasses, we developed a multilevel multivariate model. 
The detailed model is available at https://github.com/XiangZhangSC/Anitschkow.
In short, 16 subjects from the treatment group were used for this analysis, and each subject had two plasma samples at time of randomization and after treatment.   
The outcome variable is a vector of 2, $[L_i, y_i]$, in which $L_i$ represents the Lp(a) concentration in sample $i$, and $y_i$ represents the particle concentration of a lipoprotein subclass in sample $i$. 
The logarithm of $[L_i, y_i]$ was modeled by a multivariate normal distribution, with parameters $\mu_L$ (a vector of 32), $\mu_y$ (a vector of 32) and $\Sigma$ (a 2-by-2 covariance matrix). 
$\mu_L[i]$ represents the mean concentration of Lp(a) in sample $i$, and was modeled as 
$$\mu_L[i] = \alpha_{L,subject[i]} + \beta_{L,subject[i]} \times V_2[i]$$ 
$\alpha_L$ is a vector of 16, and represents the Lp(a) abundance at baseline. 
$\beta_L$ is a vector of 16, and represents the effect of evolocumab treatment on Lp(a). 
$V_2$ is a vector of 32 with 0s and 1s. 
When $V_2[i] = 1$, it means that measurement was derived from the second visit (after treatment). 
Otherwise, $V_2[i] = 0$. 
Similarly, $\mu_y[i]$ represents the mean concentration of a lipoprotein subclass in sample $i$, and was modeled as 
$$\mu_y[i] = \alpha_{y,subject[i]} + \beta_{y,subject[i]} \times V_2[i]$$ 
$\alpha_y$ is a vector of 16, and represents the lipoprotein subclass abundance at baseline. 
$\beta_y$ is a vector of 16, and represents the effect of evolocumab treatment on the lipoprotein subclass. 
To assess the relationship between the Lp(a) lowering and the metabolism of a lipoprotein subclass, we used another multivariate normal distribution to model the subject-specific parameters. 
$$[\alpha_L, \beta_L, \alpha_y, \beta_y][j] \sim MultivariateNormal([\theta_L, \gamma_L, \theta_y, \gamma_y], \Sigma_{subj})$$

$\theta_L$, $\gamma_L$, $\theta_y$, $\gamma_y$ and $\Sigma_{subj}$ (a 4-by-4 covariance matrix) are the hyper-parameters. 
We fitted the model by running Hamiltonian Markov Chain Monte Carlo in the program Stan (version 2.18.3).
We ran four Markov chains with 2000 iterations in each chain. 
Results were presented with the posterior mean with 95% credible interval (CI).

@@

The current study included 30 patients (evolocumab, n = 16; placebo, n = 14). 
Baseline characteristics of the two groups were comparable and are provided in Table 1.

@@

On average, evolocumab treatment for 16 weeks resulted in a 17% (95% credible interval: 8% to 26%) reduction in Lp(a), together with a concomitant 67% (57%, 76%) and 21% (6% 35%) reduction in LDL cholesterol and triglyceride, respectively.

To identify the metabolic effects corresponding to PCSK9 inhibition with evolocumab, we performed NMR metabolomics covering metabolic pathways such as lipoprotein subclasses, fatty acids, amino acids and glycolysis. 
We observed that evolocumab treatment resulted in substantial reduction in particle concentration of extremely large (80% [48% 100%]), very large (90% [70% 100%]), large (60% [34%, 83%]), medium (50% [36%, 63%]), small (39% [32% 46%]) and very small VLDL (47% [40% 53%]).
We also observed that evolocumab treatment resulted in particle concentration reduction in IDL (53% [45%, 60%]), large (56% [48% 65%]), medium (59% [50%, 67%]) and small LDL (55% [47%, 64%]). 
In addition, we observed that evolocumab treatment resulted in decreased concentrations of very large HDL particles (24% [3% 45%]), and increased concentrations of medium HDL particles (13% [4% 23%]) (Figure 1).

@@

Similar to the lipoprotein particle concentration profiles, we observed that evolocumab treatment resulted in substantial reduction of esterified cholesterol (CE), triglyceride (TG), free cholesterol (FC) and phospholipid (PL) in VLDL, IDL and LDL as well as very large HDL (Figure 2).

@@

We observed no difference in concentrations of other metabolites including amino acids, fluid balance, glycolysis and ketone bodies.

## Relationship between Lp(a) lowering and reduction in lipoprotein subclasses

To identify the relationship between Lp(a) lowering and reduction in lipoprotein subclasses, we developed a multilevel multivariate model. 
We observed that Lp(a) lowering was not associated with reduction of the 14 lipoprotein subclasses (Figure 4).

![Figure 4 Relationship between Lp(a) lowering and reduction in 14 lipoprotein subclasses. Circles represent the posterior mean Pearson correlation coefficients. Lines refer to the 95% credible intervals.](./corr_lipoproteins_lpa.png)

However, we identified that the reduction in medium-sized VLDL particles was associated with increasing baseline Lp(a) concentrations (Pearson correlation coefficient -0.5 [-0.8, -0.06]). 
The correlations between baseline Lp(a) concentrations and reduction in other VLDL particles, including extremely large, very large, large, small and very small VLDLs, were -0.1 [-0.6, 0.3], -0.03 [-0.4, 0.4], -0.3 [-0.7, 0.2], -0.1 [-0.7, 0.6], and 0.2 [-0.3, 0.7], respectively.

![Figure 5 Reduction of medium VLDL particles correlated with baseline lipoprotein(a) concentrations. Every circle represents the posterior mean reduction of medium VLDL particle concentration and the posterior mean of baseline lipoprotein(a) in a patient treated with evolocumab. The verticle and horizontal bar represents the 95% credible interval. The blue dashed line represented the average percentage (50%) change in medium VLDL particle after evolocumab treatment.](./Baseline_lpa_mVLDL.png)

@@

The major finding of this work is that evolocumab treatment in patients with high Lp(a) massively decreases VLDLs in addition to the well known effect on LDLs. 
In the current work, evolocumab treatment resulted in on average 17% reduction in plasma Lp(a) concentration, together with 67% and 21% reduction in LDL cholesterol and triglyceride. 
These treatment effects were similar to the outcomes reported in the past, indicating that the subjects involved in the current study were representative for the entire ANITSCHKOW trail [@Stiekema2018].

Our detailed metabolomic profiling revealed that PCSK9 inhibition with evolocumab resulted in a substantial reduction in concentrations of VLDL, IDL and LDL particles as well as their lipid and fatty acid contents. 
We did not observe changes in other circulating metabolites such as amino acids and ketone bodies. 
In general, PCSK9 inhibition with evolocumab resulted in similar metabolic changes as genetic inhibition of *PCSK9* and statin therapy [@Wurtz2016; @Sliz2018]. 
However, in contrast to genetic inhibition of *PCSK9* and statin therapy that had much larger effects on LDL particles than VLDL particles, we observed that evolocumab treatment induced a more substantial reduction in VLDL particles compared to LDL particles. 
Our multivariate analysis showed that Lp(a) lowering did not correlate with reduction in any of the 14 lipoprotein subclasses, suggesting alternative pathways in Lp(a) catabolism, such as scavenger receptor class B type I (SR-BI) [@Yang2013] and plasminogen receptor [@Sharma2017].

Interestingly, we identified that subjects with higher baseline Lp(a) showed a larger reduction in medium VLDL particle concentrations after treatment with evolocumab. 
We observed that subjects with higher baseline Lp(a) also showed tendency to have a larger reduction in extremely large and large VLDL particle concentrations after treatment with evolocumab. 
However, those tendencies had large uncertainty due to considerable number of missing observations. 
The percentage of missing observations in extremely large and large were 69% and 47%, respectively. 
Similarly, since the very large VLDL particle had the largest percentage of missing data (72%) among all the 14 lipoprotein subclasses, we simply had not enough data to detect the relationship between its reduction and baseline Lp(a).  
In contrast, the percentage of missing observations in medium VLDL was 9%, allowing us to assess the relationship between its reduction and baseline Lp(a) levels. 
On the other hand, we did not observe similar relationship between baseline Lp(a) and reduction in small and very small VLDL particles, although the percentages of missing observations were small for these two lipoprotein subclasses (6% and 3%), suggesting that Lp(a) may not influence catabolism of these two triglyceride-poor VLDL particles.    
Our observation is in line with the association between the Lp(a)-raising genotype and decreasing concentrations of extremely large, very large, large and medium VLDL particles, but not small and very small VLDL particles [@Kettunen2016]. 
Lp(a) particles can bind noncovalently to triglyceride-rich lipoproteins (TRL), forming a Lp(a)-TRL complex which was suggested to facilitate receptor-mediated uptake [@McConathy1994; @Gaubatz2001]. 
This process may be particularly important in our subjects with high Lp(a), since they may have a large amount of the Lp(a)-TRL complex. 
Together with the supraphysiological abundance of lipoprotein receptors induced by evolocumab treatment, these two factors may underlie the massive reduction of VLDL particles. 
On the other hand, apo(a) can be recycled to the extracellular space and bind to apoB-containing particles again, resulting in moderate reduction of Lp(a) [@Sharma2017]. 
We find that our hypothesis has connections to other studies. 
First, in the ANITSCHKOW trail, evolocumab had no impact on arterial wall inflammation in patients with persistent Lp(a) elevation [@Stiekema2018]. 
This could be caused by the fact that inhibition of PCSK9 may increase other receptors such as VLDL receptors and CD36. 
Both the VLDL receptor and CD36 are present in macrophages and endothelial cells. 
The accelerated uptake of a Lp(a)-TRL complex via these receptors would not attenuate foam cell formation. 
Second, when Lp(a) is directly targeted via antisense oligonucleotide treatment, reduction of Lp(a) production does decrease cellular inflammatory responses [@Viney2016]. 
In this case the reduced Lp(a) production may decrease the abundance of Lp(a)-TRL complex and uptake of TRL by macrophages, thereby reducing foam cell formation. 
Future studies are required in order to test our hypothesis.

@@

Our study provided the first NMR metabolomics data for a PCSK9 inhibition trial. 
The detailed metabolomic profiling not only allowed us to reveal systemic effects of PCSK9 inhibition, but also to assess the relationship between Lp(a) lowering and 14 lipoprotein subclasses. 
We did not detect statistically significant correlations between Lp(a) lowering and reduction in lipoprotein subclasses probably due to the small number of participants and considerable missing data in some lipoprotein subclasses. 
We feel it is worth to apply the multilevel multivariate model developed in this study in a larger clinical trial to improve our understanding of interrelationship between Lp(a) lowering and metabolism of other lipoprotein subclasses.

In conclusion, our NMR metabolomic profiling revealed that evolocumab treatment in patients with high Lp(a) massively decreases VLDLs in addition to the well known effect on LDLs. 
The substantial reduction in VLDLs after evolocumab treatment suggested enhanced clearance of VLDL particles in patients with high Lp(a).
